Probabilistic Catalogs from Crowded Star Fields

### 2012-07-11 paper/ms.tex
Bayesian waffling

\section{The Unknown Parameters}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the {\it spatial distribution} of the stars by parameters
$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
function}) by parameters $\beta$. In summary, the unknown parameters are:
\begin{eqnarray}
\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
p(f_i | \beta) 
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
others. Finally, the positions and fluxes of the stars come independently from
a common distribution. If we knew the spatial distribution and the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star.

\section{Simulated Data}

>>>>

\section{Bayesian Inference}
We use Bayesian Inference. Suppose there exist unknown parameters (denoted
collectively by $\theta$) and we expect to obtain some data $D$. Our prior
knowledge (or uncertainty) about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta)
\end{equation}
We also model how the parameters give rise to the data, via a {\it sampling
distribution}:
\begin{equation}
p(D|\theta)
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions, one for each possible value of $\theta$. Note that
the sampling distribution is also an assertion of a prior state of knowledge.
In this case it models our prior information about the fact that the parameters
$\theta$ are going to influence the data in some way (otherwise this data
would be irrelevant).

When specific data $D^*$ are known, the prior gets updated to the posterior
via Bayes' rule:
\begin{eqnarray}
p(\theta|D=D^*) &\propto& p(\theta)p(D|\theta)|_{D=D^*} \\
&=& p(\theta)\mathcal{L}(\theta; D)
\end{eqnarray}
The modulating function is the {\it likelihood function}, which is the prior
probability of obtaining the actual data set $D^*$ as a function of the
parameters. As suggested by the above notation, it is obtained from the
sampling distribution with the actual data substituted in.

\section{The Specific Model for Stellar Fields}

\subsection{The Unknown Parameters}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the {\it spatial distribution} of the stars by parameters
$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
function}) by parameters $\beta$. In summary, the unknown parameters are:
\begin{eqnarray}
\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
p(f_i | \beta) 
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
others. Finally, the positions and fluxes of the stars come independently from
a common distribution. If we knew the spatial distribution and the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star.

\subsection{The Sampling Distribution}

\subsection{The Prior Distribution}

{Simulated Data}
